---
title: "Gathering FP Data for Analysis"
author: "João Martins"
output: 
  html_document: 
    theme: cerulean
    code_folding: show
    toc: true
    toc_float: true
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r boilerplate, message=FALSE, warning=FALSE, comment = FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(glue)
library(readxl)
library(vroom)
library(yaml)
library(httr)
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = FALSE)
source(here("R", "utils.R"))
```

## Data Sources

The raw data is available at:

- [CORDIS](https://cordis.europa.eu/)
  - [CORDIS FP7 projects](https://data.europa.eu/data/datasets/cordisfp7projects)  
  - [CORDIS H2020 projects](https://data.europa.eu/data/datasets/cordish2020projects), updated 2021-12-14
  - [CORDIS reference data](https://data.europa.eu/data/datasets/cordisref-data)
  - [Country data]()
- [**Horizon Dashboard**](https://webgate.ec.europa.eu/dashboard) 
 - `Everyone` ➡ `FP7 Projects` ➡ `FP7 Project Details` ➡ `Top Funded Projects`  
 - `Everyone` ➡ `H2020 Projects` ➡ `H2020 Projects` ➡ `Top Funded Projects`  
 - `Everyone` ➡ `H2020 Projects` ➡ `Data Export` ➡ `Raw Data Export Sheet`, updated 2021-12-06  

> Horizon Europe project details have not yet been published.

## Collecting CORDIS Data

CORDIS publishes snapshots of its public data in the [EU open data portal](https://data.europa.eu).  The data are provided by framework programme.  For FP7 and H2020, the latest data are now packed in ZIP files containing a collection of CSV datasets.  We can see the details of how to download them in the code.

```{r extract-download-cordis-data, eval=FALSE}
end_points <-
  expand_grid(c("fp6", "fp7", "h2020"),
              c("projects", "organizations")) %>%
  setNames(c("programme",
             "dataset")) %>%
  mutate(dataset_name = sprintf("cordis-%s%s.csv", programme, dataset)) %>%
  pull(dataset_name)

# updated 09 Mar 2020 ----
glue("https://cordis.europa.eu/data/FP6/",
     "{end_points[1]}") %>%
  httr::HEAD() %>%
  httr::headers() %>%
  pluck("last-modified")
glue("https://cordis.europa.eu/data/FP6/",
     "{end_points[2]}") %>%
  httr::HEAD() %>%
  httr::headers() %>%
  pluck("last-modified")

glue("https://cordis.europa.eu/data/FP6/",
     "{end_points[1]}") %>%
  download.file(here("data-raw", end_points[1]))
glue("https://cordis.europa.eu/data/FP6/",
     "{end_points[2]}") %>%
  download.file(here("data-raw", end_points[2]))

# updated 17 Aug 2021 ----
glue("https://cordis.europa.eu/data/",
     "{end_points[3]}") %>%
  httr::HEAD() %>%
  httr::headers() %>%
  pluck("last-modified")
glue("https://cordis.europa.eu/data/",
     "{end_points[4]}") %>%
  httr::HEAD() %>%
  httr::headers() %>%
  pluck("last-modified")
glue("https://cordis.europa.eu/data/",
     "{end_points[5]}") %>%
  httr::HEAD() %>%
  httr::headers() %>%
  pluck("last-modified")
glue("https://cordis.europa.eu/data/",
     "{end_points[6]}") %>%
  httr::HEAD() %>%
  httr::headers() %>%
  pluck("last-modified")

glue("https://cordis.europa.eu/data/",
     "{end_points[3]}") %>%
  download.file(here("data-raw", end_points[3]))
glue("https://cordis.europa.eu/data/",
     "{end_points[4]}") %>%
  download.file(here("data-raw", end_points[4]))
glue("https://cordis.europa.eu/data/",
     "{end_points[5]}") %>%
  download.file(here("data-raw", end_points[5]))
glue("https://cordis.europa.eu/data/",
     "{end_points[6]}") %>%
  download.file(here("data-raw", end_points[6]))

# updated 18 Jan 2022 ----
"https://cordis.europa.eu/data/cordis-fp7projects-csv.zip" %>%
  httr::HEAD() %>%
  httr::headers() %>%
  pluck("last-modified")
"https://cordis.europa.eu/data/cordis-h2020projects-csv.zip" %>%
  httr::HEAD() %>%
  httr::headers() %>%
  pluck("last-modified")
"https://cordis.europa.eu/data/cordis-fp7projects-csv.zip" %>%
  download.file(here("data-raw", "cordis-fp7projects-csv.zip"))
"https://cordis.europa.eu/data/cordis-h2020projects-csv.zip" %>%
  download.file(here("data-raw", "cordis-h2020projects-csv.zip"))
unzip(
  here("data-raw", "cordis-fp7projects-csv.zip"),
  exdir = here("data-raw", "fp7")
)
unzip(
  here("data-raw", "cordis-h2020projects-csv.zip"),
  exdir = here("data-raw", "h2020")
)

# clanup
rm(end_points)
```

## Collecting Dashboard Data

Horizon Dashboard is built on Qlik Sense which prevents users from downloading the raw data automatically. The data can nevertheless be exported manually.

### Gathering Projects

```{r load-projects-data}
# Dashboard data (FP7, H2020 only) ----
h2020_dashboard <- here("data-raw", "h2020-projects.xlsx") %>%
  read_excel(guess_max = 20000) %>%
  clean_names() %>%
  select(
    project_id = project_nbr,
    project_acronym,
    thematic_priority = thema,
    topic_code,
    topic_description,
    participations = h2020_participations) %>%
  mutate(framework_programme = "H2020")

fp7_dashboard <- here("data-raw", "fp7-projects.xlsx") %>%
  read_excel(guess_max = 20000) %>%
  clean_names() %>%
  select(
    project_id = project_nbr,
    project_acronym,
    thematic_priority = thematic_priority_descr,
    topic_code,
    topic_description,
    participations) %>%
  mutate(framework_programme = "FP7")

projects_dashboard <-
  fp7_dashboard %>%
  bind_rows(h2020_dashboard)

# CORDIS data ----

# reconcile schemas (using latest schema as reference )
projects_cordis_h2020 <- 
  read_csv2(here("data-raw", "h2020", "csv", "project.csv")) %>%
  mutate(contentUpdateDate = as.Date(contentUpdateDate))

projects_cordis_fp7 <-
  readr::read_csv2(here("data-raw", "fp7", "csv", "project.csv")) %>%
  mutate(
    contentUpdateDate = as.Date(contentUpdateDate),
    ecSignatureDate = as.Date(ecSignatureDate), # missing
    masterCall = as.character(masterCall)) # missing

projects_cordis <-
  projects_cordis_fp7 %>%
  bind_rows(projects_cordis_h2020) %>%
  clean_names() %>%
  select(
    project_id = id,
    start_date,
    end_date,
    legal_basis,
    funding_scheme)

projects <-
  projects_dashboard %>%
  left_join(projects_cordis, by = "project_id")

# cleanup
rm(
  fp7_dashboard,
  h2020_dashboard,
  projects_dashboard,
  projects_cordis,
  projects_cordis_fp7,
  projects_cordis_h2020
)
```

Unfortunately, FP6 data are very incomplete. Individual contribution details are provided in 2,621 (out of 75,241) cases only.

```{r fp6, eval=FALSE}
projects_cordis_fp6 <-
  readr::read_csv2(here("data-raw", "cordis-fp6projects.csv")) %>%
  clean_names() %>%
  select(
    project_id = id,
    rcn,
    proj_acronym = acronym,
    start_date,
    end_date,
    legal_basis = call,
    framework_programme,
    funding_scheme
  )

organizations_cordis_fp6 <-
  read_tsv(here("data-raw", "cordis-fp6organizations.csv"))
```

### Gathering Participations

```{r, load-organizations-data}
organizations_cordis_h2020 <- 
  read_csv2(here("data-raw", "h2020", "csv", "organization.csv"), guess_max = 100000) %>%
  select(-contentUpdateDate, -SME, -nutsCode, -endOfParticipation, -active, -totalCost, -rcn)

organizations_cordis_fp7 <- 
  read_csv2(here("data-raw", "fp7", "csv", "organization.csv"), guess_max = 130000) %>%
  select(-contentUpdateDate, -SME, -nutsCode, -endOfParticipation, -active, -totalCost, -rcn)
  
organizations_cordis <-
  organizations_cordis_fp7 %>%
  bind_rows(organizations_cordis_h2020) %>%
  clean_names() %>%
  select(
    project_id,
    pic = organisation_id,
    legal_name = name,
    legal_short_name = short_name,
    legal_entity_type = activity_type,
    country_code = country,
    legal_url = organization_url,
    partner_role = role,
    ec_contribution,
    net_ec_contribution)
  # ! 3 cases where (project_id, pic) appear 2x w/ different partner_role

organizations_dashboard_h2020 <-
  here("data-raw", "horizon2020.xlsx") %>%
  read_excel() %>%
  clean_names() %>%
  select(
    project_id = project_nbr,
    pic = general_pic,
    partner_role,
    pillar_abbr,
    pillar_descr,
    thematic_priority_abbr,
    signature_date,
    call_deadline_date,
    eu_contribution
  ) %>%
  mutate(partner_role = str_to_lower(partner_role))

organizations <- organizations_cordis %>%
  left_join(organizations_dashboard_h2020,
    by = c("project_id", "pic", "partner_role"))

rm(organizations_cordis, organizations_dashboard_h2020)
```

```{r viz-contributions, eval=FALSE}
dfsample <- organizations %>%
  filter(!is.na(net_ec_contribution)) %>%
  sample_n(10000)
plot(
  dfsample$ec_contribution,
  dfsample$eu_contribution,
  type = "p",
  pch = 19,
  cex = .5,
  log = 'xy'
)
plot(
  dfsample$net_ec_contribution,
  dfsample$eu_contribution,
  type = "p",
  pch = 19,
  cex = .5,
  log = 'xy'
)
plot(
  dfsample$net_ec_contribution,
  dfsample$ec_contribution,
  type = "p",
  pch = 19,
  cex = .5,
  log = 'xy'
)

```

### Combining project and participation data

```{r transform}
df_raw <- organizations %>%
  left_join(projects, by = "project_id")

# download CORDIS country codes
country_codes <-
  "https://cordis.europa.eu/data/reference/cordisref-countries.csv" %>%
  read_csv2() %>%
  clean_names() %>%
  filter(language == "en") %>%
  distinct(eu_code, name) %>%
  rename(country_code = eu_code, country = name)

df_raw <- df_raw %>%
  mutate(
    legal_entity_type = str_sub(legal_entity_type, 1, 3),
    country_code = str_sub(country_code, 1, 2),
    # rename role
    partner_role = str_to_title(partner_role),
    pillar = str_extract(pillar_abbr, "\\d"),
    pillar = if_else(pillar_abbr == "EU.0.", "Cross-theme", pillar),
    pillar = if_else(pillar_abbr == "Euratom", pillar_abbr, pillar),
    # clean contributions
    ec_contribution = str_replace(ec_contribution, ",", "."),
    ec_contribution = str_replace(ec_contribution,
                                  fixed("xxxxx"), 
                                  NA_character_),
    eu_contribution = str_replace(eu_contribution, fixed("-"), "0"),
    ec_contribution = as.numeric(ec_contribution),
    net_ec_contribution = as.numeric(net_ec_contribution),
    eu_contribution = as.numeric(eu_contribution),
    # reconcile FP7 and H2020 contribution data
    re_contribution = eu_contribution,
    re_contribution = if_else(is.na(re_contribution),
                              net_ec_contribution,
                              re_contribution),
    re_contribution = if_else(is.na(re_contribution),
                              ec_contribution,
                              re_contribution),
    # country code
    country_code = str_replace(country_code, "KO", "XK"),
    # import dates
    signature_date = as.Date(signature_date, format = "%d/%m/%Y"),
    call_deadline_date = as.Date(call_deadline_date, format = "%d/%m/%Y"),
    signature_year = str_sub(signature_date, 1, 4),
    call_year = str_sub(call_deadline_date, 1, 4),
    start_year = str_sub(start_date, 1, 4),
    signature_year = int(signature_year),
    call_year = int(call_year),
    start_year = int(start_year)
  ) %>%
  left_join(country_codes, by = "country_code") %>%
  filter(!is.na(start_date))

```

## Exporting data for analysis

```{r write}
# store data types
df_schema <- df_raw %>%
  map_chr(class)

# split CSV by year
df_years <-
  df_raw %>%
  distinct(start_year) %>%
  pull() %>%
  sort()
df_raw <- df_raw %>%
  arrange(start_year) %>%
  group_by(start_year) %>%
  group_split() %>%
  setNames(df_years)

# write to ./data/
file_names <- here("data", glue("cordis-plus-{df_years}.csv"))
walk2(df_raw, file_names, vroom_write, delim = ",")
write_yaml(df_schema, here("data", "schema-cordis-plus.yml"))

```

