---
title: "Reproducibility analylsis of the SERI figures"
subtitle: "On the Swiss Participation in the EU-Framework Programmes for Research and Innovation"
author: "João Martins"
output: 
  html_document: 
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r boilerplate, message=FALSE, warning=FALSE, comment = FALSE}
# source code: https://github.com/zambujo/swissparticipation
library(tidyverse)
library(vroom)
library(janitor)
library(lubridate)
library(here)
library(yaml)
library(glue)
library(scales)
library(ggsci)
library(ggrepel)
library(ggtext)
library(kableExtra)
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = FALSE)
source(here("R", "utils.R"))
```

> Pease find related statistics on the SERI's page: [Facts and Figures on the Swiss Participation in the EU-Framework Programmes for Research and Innovation](https://www.sbfi.admin.ch/sbfi/en/home/research-and-innovation/international-cooperation-r-and-i/eu-framework-programmes-for-research/f-f-swiss-participation.html).

```{r read-dashboard-data}
df_schema <- here("data", "schema-cordis-plus.yml") %>%
  read_yaml() %>%
  as.list()

df_raw <-
  list.files(here("data"), pattern = "cordis-plus-", full.names = TRUE) %>%
  vroom(delim = ",", col_types = df_schema)
  # map_df(read_csv2, col_types = df_schema)

df_raw <-
  df_raw %>%
  mutate(
    pillar = fct_inseq(pillar),
    pillar = fct_rev(pillar),
    pillar = fct_relevel(pillar, "Cross-theme", after = 0),
    pillar = fct_recode(
      pillar, I = "1", II = "2", III = "3", IV = "4", V = "5"),
    org_type = fct_collapse(
      legal_entity_type,
      `Research & Education` = c("HES", "REC"),
      `Others (Excl. R&Ed)` = c("PRC", "PUB", "OTH")))

df <- df_raw %>%
    mutate(start_semester = floor_date(start_date, "6 months")) %>%
    # remove the CERN
    filter(legal_name != "EUROPEAN ORGANIZATION FOR NUCLEAR RESEARCH")
```

## Overview  {.tabset .tabset-fade .tabset-pills}

```{r plot-styling}
bar_plt <- function(plt) {
  plt +
  geom_col() +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    text = element_text(
      size = 15,
      family = "Gill Sans",
      color = "gray15"
    )
  )
}

get_particip <- function(x, in_programme = "") {
  if (in_programme == "FP7") {
    x <- x %>%
      filter(framework_programme == "FP7")
  }
  if (in_programme == "H2020") {
    x <- x %>%
      filter(framework_programme == "H2020")
  }
  total <- x %>%
    count(Semester = start_semester, name = "n_total")
  ch <- x %>%
    filter(country_code == "CH") %>%
    count(Semester = start_semester, name = "n_ch")
  
  left_join(total, ch, by = "Semester") %>%
    mutate(year = year(Semester),
           share_particip = 100 * n_ch / n_total)
}

get_contrib <- function(x, in_programme = "") {
  if (in_programme == "FP7") {
    x <- x %>%
      filter(framework_programme == "FP7")
  }
  if (in_programme == "H2020") {
    x <- x %>%
      filter(framework_programme == "H2020")
  }
  to_all <- x %>%
    group_by(Semester = start_semester) %>%
    summarise(contrib_all = sum(re_contribution, na.rm = TRUE))
  to_ch <- x %>%
    filter(country_code == "CH") %>%
    group_by(Semester = start_semester) %>%
    summarise(contrib_ch = sum(re_contribution, na.rm = TRUE))

  left_join(to_all, to_ch, by = "Semester") %>%
    mutate(year = year(Semester),
           share_contrib = 100 * contrib_ch / contrib_all)
}

get_coord <- function(x, in_programme = "") {
  if (in_programme == "FP7") {
    x <- x %>%
      filter(framework_programme == "FP7")
  }
  if (in_programme == "H2020") {
    x <- x %>%
      filter(framework_programme == "H2020")
  }
  
  all <- x %>%
    filter(partner_role == "Coordinator") %>%
    count(Semester = start_semester, name = "coord_total")
  ch <- x %>%
    filter(country_code == "CH") %>%
    filter(partner_role == "Coordinator") %>%
    count(Semester = start_semester, name = "coord_ch")

  left_join(all, ch, by = "Semester") %>%
    mutate(year = year(Semester),
           share_contrib = 100 * coord_ch / coord_total)
}
```

SERI figures show slightly higher shares of contributions to Swiss participants. They should, nevertheless, exclude the CERN, the UN, and UN-affiliated institutions.

### Share Swiss Participations

```{r swiss-participations}
df %>%
  get_particip() %>%
  ggplot(aes(x = year, y = share_particip)) %>%
  bar_plt() +
  labs(
    title = "Swiss Participation",
    subtitle = "By project starting year, FP7 and H2020",
    x = "Project Starting Year",
    y = "%"
  ) +
  annotate(
    "rect",
    xmin = 2006.5,
    xmax = 2014.5,
    ymin = -.1,
    ymax = 8.1,
    fill = "#21325E",
    alpha = .1
  ) +
  annotate(
    "rect",
    xmin = 2014.5,
    xmax = 2022.5,
    ymin = -.1,
    ymax = 8.1,
    fill = "#F1D00A",
    alpha = .1
  ) +
  annotate(
    "text",
    x = c(2010, 2019),
    y = c(8.4, 8.4),
    label = c("FP7", "H2020"),
    family = "Gill Sans"
  )
```

```{r swiss-participations-fp7, include=FALSE}
df %>%
  get_particip(in_programme = "FP7") %>%
  select(n_ch, n_total) %>%
  summarise_all(sum, na.rm = TRUE) %>%
  mutate(share = round(100 * n_ch / n_total, 1)) %>%
  kbl()
```

```{r swiss-participations-h2020, include=FALSE}
df %>%
  get_particip(in_programme = "H2020") %>%
  select(n_ch, n_total) %>%
  summarise_all(sum, na.rm = TRUE) %>%
  mutate(share = round(100 * n_ch / n_total, 1)) %>%
  kbl()
```

### Contribution to Swiss Participants

```{r swiss-countributions}
df %>%
  get_contrib() %>%
  ggplot(aes(x = year, y = share_contrib)) %>%
  bar_plt() +
  labs(
    title = "Contributions to Swiss Participants",
    subtitle = "By project starting year, FP7 and H2020",
    x = "Project Starting Year",
    y = "%"
  ) +
  annotate(
    "rect",
    xmin = 2006.5,
    xmax = 2014.5,
    ymin = -.1,
    ymax = 12.7,
    fill = "#21325E",
    alpha = .1
  ) +
  annotate(
    "rect",
    xmin = 2014.5,
    xmax = 2022.5,
    ymin = -.1,
    ymax = 12.7,
    fill = "#F1D00A",
    alpha = .1
  ) +
  annotate(
    "text",
    x = c(2010, 2019),
    y = c(13.1, 13.1),
    label = c("FP7", "H2020"),
    family = "Gill Sans"
  )
```

```{r contributions-ch-fp7, include=FALSE}
df %>%
  get_contrib(in_programme = "FP7") %>%
  select(contrib_ch, contrib_all) %>%
  summarise_all(sum, na.rm = TRUE) %>%
  mutate(share = round(100 * contrib_ch / contrib_all, 1)) %>%
  kbl()
```

```{r contributions-ch-h2020, include=FALSE}
df %>%
  get_contrib(in_programme = "H2020") %>%
  select(contrib_ch, contrib_all) %>%
  summarise_all(sum, na.rm = TRUE) %>%
  mutate(share = round(100 * contrib_ch / contrib_all, 1)) %>%
  kbl()
```

### Swiss Coordinations

```{r coordinations-overviews}
df %>%
  get_coord() %>%
  ggplot(aes(x = year, y = share_contrib)) %>%
  bar_plt() +
  labs(
    title = "Swiss Coordinations",
    subtitle = "By project starting year, FP7 and H2020",
    x = "Project Starting Year",
    y = "%"
  ) +
  annotate(
    "rect",
    xmin = 2006.5,
    xmax = 2014.5,
    ymin = -.1,
    ymax = 11.1,
    fill = "#21325E",
    alpha = .1
  ) +
  annotate(
    "rect",
    xmin = 2014.5,
    xmax = 2022.5,
    ymin = -.1,
    ymax = 11.1,
    fill = "#F1D00A",
    alpha = .1
  ) +
  annotate(
    "text",
    x = c(2010, 2019),
    y = c(11.4, 11.4),
    label = c("FP7", "H2020"),
    family = "Gill Sans"
  )
```

```{r coordinations-ch-fp7, include=FALSE}
df %>%
  get_coord(in_programme = "FP7") %>%
  select(coord_ch, coord_total) %>%
  summarise_all(sum, na.rm = TRUE) %>%
  mutate(share = round(100 * coord_ch / coord_total, 1)) %>%
  kbl()
```

```{r coordinations-ch-h2020, include=FALSE}
df %>%
  get_coord(in_programme = "H2020") %>%
  select(coord_ch, coord_total) %>%
  summarise_all(sum, na.rm = TRUE) %>%
  mutate(share = round(100 * coord_ch / coord_total, 1)) %>%
  kbl()
```

## H2020 Figures {.tabset .tabset-fade .tabset-pills}

The below *Figure 4* updates [Figure 17 in page 30](https://www.sbfi.admin.ch/dam/sbfi/en/dokumente/webshop/2018/ch-frp.pdf.download.pdf/ch_beteiligung_frp_e.pdf#page=30). *Figure 4 (Original)* corresponds to the latest update provided in the [SERI page](https://www.sbfi.admin.ch/sbfi/en/home/research-and-innovation/international-cooperation-r-and-i/eu-framework-programmes-for-research/f-f-swiss-participation.html). *Figure 4 (Updated)* shows the latest H2020 dashboard data.  (Only the 20 top participation areas are shown.)

### Figure 4 (Original)

![](https://www.sbfi.admin.ch/sbfi/en/home/research-and-innovation/international-cooperation-r-and-i/eu-framework-programmes-for-research/f-f-swiss-participation/_jcr_content/par/accordion_16708559/items/abbildung_4_beitr_ge/accordionpar/image_896504598/image.imagespooler.png/1629209007946/abb4_en.png "SERI Figure 4. Funding under by programme area (CHF m), in relation to total funding")

### Figure 4 (Updated)

```{r fig-four, fig.asp=1.5}
get_contrib_area <- function(x) {
  x <- x %>% 
    filter(framework_programme == "H2020") %>%
    filter(!is.na(thematic_priority_abbr))

  to_all <- x %>%
    group_by(Area = thematic_priority_abbr) %>%
    summarise(contrib = sum(eu_contribution, na.rm = TRUE)) %>%
    mutate(
      share = contrib / sum(contrib),
      context = "Total")
  to_ch <- x %>%
    filter(country_code == "CH") %>%
    group_by(Area = thematic_priority_abbr) %>%
    summarise(contrib = sum(eu_contribution, na.rm = TRUE)) %>%
    mutate(
      share = contrib / sum(contrib),
      context = "CH")
  bind_rows(to_all, to_ch)
}

pal <- c("#B91646", "#DFD8CA") %>% rev()

pillar_area_lookup <- df %>%
  filter(framework_programme == "H2020") %>%
  count(Pillar = pillar, Area = thematic_priority_abbr) %>%
  filter(!is.na(Area)) %>%
  select(-n)

area_labels <- df %>%
  get_contrib_area() %>%
  filter(context == "CH") %>%
  left_join(pillar_area_lookup, by = "Area") %>%
  arrange(desc(contrib)) %>%
  slice_head(n = 20) %>%
  arrange(desc(Pillar), desc(contrib)) %>%
  select(Area)

contribs <- df %>%
  get_contrib_area() %>%
  select(-share) %>%
  mutate(
    contrib = round(1e-6 * contrib, 1),
    contrib = format(contrib, big.mark = ",", decimal.mark = ".", trim = TRUE))

area_labels %>%
  left_join(get_contrib_area(df), by = "Area") %>%
  select(-contrib) %>%
  pivot_wider(names_from = context, values_from = share) %>%
  pivot_longer(cols = c("Total", "CH"), names_to = "context", values_to = "share") %>%
  left_join(contribs, by = c("Area", "context")) %>%
  mutate(
    context = str_replace(context, "CH", "Switzerland"),
    Area = str_remove(Area, "LEIT-"),
    Area = str_remove(Area, "INNOSUP"),
    Area = fct_inorder(Area),
    Area = fct_rev(Area)) %>%
  ggplot(aes(x = Area, y = share, fill = fct_rev(context))) +
  geom_bar(stat = "identity", width = .7, position = "dodge") +
  geom_text(
    aes(label = contrib),
    position = position_dodge(width = .9),
    hjust = -.25,
    size = 3.5,
    family = "Gill Sans"
  ) +
  coord_flip() +
  scale_fill_manual(values = pal, guide = guide_legend(reverse = TRUE)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), expand = c(.1, 0)) +
  theme_minimal() +
  labs(
    title = "Funding by Programme Area Under H2020",
    subtitle = "Values in MioEUR")+
  theme(
    text = element_text(size = 14, family = "Gill Sans", color = "gray10"),
    plot.title.position = "plot",
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_text(margin = margin(0,-25,0,0)))
```
