---
title: "Swiss Participation in Horizon 2020"
output: 
  html_document:
    self_contained: no
---

```{r boilerplate, include=FALSE}
knitr::opts_chunk$set(
  dev = "svg",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = FALSE
)
```

```{r read-dashboard-data}
source(here::here("R", "package.r"))
source(here::here("R", "utils.r"))

df_schema <- here("data", "schema-cordis-plus.yml") %>%
  read_yaml() %>%
  as.list()

df_raw <-
  list.files(here("data"), pattern = "cordis-plus-", full.names = TRUE) %>%
  map_df(read_csv, col_types = df_schema)

df_raw <- df_raw %>%
  mutate(
    pillar = fct_inseq(pillar),
    pillar = fct_rev(pillar),
    pillar = fct_relevel(pillar, "Cross-theme", after = 0),
    pillar = fct_recode(
      pillar, I = "1", II = "2", III = "3", IV = "4", V = "5"),
    org_type = fct_collapse(
      legal_entity_type,
      `Research & Education` = c("HES", "REC"),
      `Others (Excl. R&Ed)` = c("PRC", "PUB", "OTH")))

df <- df_raw %>%
  filter(framework_programme == "H2020") %>%
  filter(legal_name != "EUROPEAN ORGANIZATION FOR NUCLEAR RESEARCH")
```

## Total EU Contribution by Pillar

```{r overview-pillar, fig.asp=.7}
budget_all <- df %>%
  mutate(pillar = fct_other(pillar, keep = chr(roman(1:5))),
         pillar = fct_relevel(pillar, "Other", after = 0)) %>%
  filter(!is.na(pillar)) %>%
  group_by(pillar) %>%
  summarise(size_pillar = sum(eu_contribution)) %>%
  mutate(
    share = 100 * prop.table(size_pillar),
    fp = "H2020",
    cube = "All Countries"
  )

budget_ch <- df %>%
  filter(country_code == "CH") %>%
  mutate(pillar = fct_other(pillar, keep = chr(roman(1:5))),
         pillar = fct_relevel(pillar, "Other", after = 0)) %>%
  filter(!is.na(pillar)) %>%
  group_by(pillar) %>%
  summarise(size_pillar = sum(eu_contribution)) %>%
  mutate(
    share = 100 * prop.table(size_pillar),
    fp = "H2020",
    cube = "Switzerland"
  )

pal <- c("#5298BD",
         "#FBBE5E",
         "#71B294",
         "#9D90B9",
         "#F08262",
         "#83D0F5")

budget_ch %>%
  bind_rows(budget_all) %>%
  mutate(
    cube = factor(cube),
    pillar = fct_recode(
      pillar,
      `I Excellent Science` = "I",
      `II Insdustrial Leadership` = "II",
      `III Societal Challenges` = "III",
      `IV Spreading Excellence and Widening Participation` = "IV",
      `V Science with and for Society` = "V",
      `Euratom + Cross-theme` = "Other"
    )
  ) %>%
  ggplot(aes(x = fp, y = share, fill = pillar)) +
  geom_col(position = position_stack()) +
  geom_text_repel(
    aes(label = sprintf("%.1f%%", share)),
    position = position_stack(vjust = .5, ),
    size = 4,
    alpha = 1,
    show.legend = FALSE,
    box.padding = .75,
    family = "Gill Sans"
  ) +
  coord_flip() +
  facet_wrap(cube ~ ., ncol = 1) +
  scale_fill_manual(
    values = rev(pal),
    guide = guide_legend(ncol = 2, reverse = TRUE)) +
  labs(x = NULL, y = NULL) +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.x = element_text(size = 15),
    legend.title = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    text = element_text(
      size = 14,
      family = "Gill Sans",
      color = "gray10"
    )
  )
```


```{r top-participants-for-ref}
# find and sort top participants
top_participants <- df %>%
  group_by(Country = country) %>%
  summarise(contrib_country = sum(eu_contribution, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(contrib_country))

ch_rank <- which(pull(top_participants, Country) == "Switzerland")

# calculate further participation statistics
country_participation <- df %>%
  group_by(
    Country = country,
    `Call Year` = call_year,
    Type = org_type
  ) %>%
  summarise(
    `Participation Count` = n(),
    `Nb Projects` = n_distinct(project_id),
    `EU Contribution` = sum(eu_contribution, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(`Type`, `Call Year`)
```

## Overview

```{r get-participation}
get_participation <- function(x) {
  x %>%
    filter(!is.na(pillar)) %>%
    mutate(pillar = fct_rev(pillar)) %>%
    group_by(
      Pillar = pillar,
      Priority = thematic_priority_abbr,
      `Call Year` = call_year
    ) %>%
    summarise(
      `Participation Count` = n(),
      `Nb Projects` = n_distinct(project_id),
      `EU Contribution` = sum(eu_contribution, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    mutate(`EU Contribution` = round(`EU Contribution`, 1)) %>%
    reactable(
      groupBy = c("Pillar", "Priority"),
      columns = list(
        Pillar = colDef(footer = "Total"),
        `Participation Count` = colDef(
          aggregate = "sum",
          format = colFormat(separators = TRUE),
          footer = function(values) {
            sum(values) %>% format(big.mark = ",")
          }
        ),
        `Nb Projects` = colDef(
          aggregate = "sum",
          format = colFormat(separators = TRUE),
          footer = function(values) {
            sum(values) %>% format(big.mark = ",")
          }
        ),
        `EU Contribution` = colDef(
          aggregate = "sum",
          format = colFormat(
            currency = "EUR",
            separators = TRUE,
            digits = 2
          ),
          footer = function(values) {
            sum(values) %>%
              format(big.mark = ",") %>%
              str_c("€", .)
          }
        )
      ),
      defaultColDef = colDef(footerStyle = list(fontWeight = "bold")),
      highlight = TRUE,
      compact = TRUE
    )
}
```

Swiss participation in H2020: participation count, number of projects (distinct), total EU contribution by pillar, thematic priority, and call year.

```{r overview}
df %>%
  filter(country_code == "CH") %>%
  get_participation()
```

## Areas (Thematic Priorities) {.tabset .tabset-fade .tabset-pills}

The below *Figure 4* updates [Figure 17 in page 30](https://www.sbfi.admin.ch/dam/sbfi/en/dokumente/webshop/2018/ch-frp.pdf.download.pdf/ch_beteiligung_frp_e.pdf#page=30). **Figure 4 (Original)** is the latest update provided in the [SERI page](https://www.sbfi.admin.ch/sbfi/en/home/research-and-innovation/international-cooperation-r-and-i/eu-framework-programmes-for-research/f-f-swiss-participation.html). **Figure 4 (Updated)** shows the latest H2020 dashboard data. (For comparison, only the 20 top participation areas are shown.)

### Figure 4 (Updated)

```{r fig-four, fig.asp=1}
get_contrib_area <- function(x) {
  x <- x %>% 
    filter(framework_programme == "H2020") %>%
    filter(!is.na(thematic_priority_abbr))

  to_all <- x %>%
    group_by(Area = thematic_priority_abbr) %>%
    summarise(contrib = sum(eu_contribution, na.rm = TRUE)) %>%
    mutate(
      share = contrib / sum(contrib),
      context = "Total")
  to_ch <- x %>%
    filter(country_code == "CH") %>%
    group_by(Area = thematic_priority_abbr) %>%
    summarise(contrib = sum(eu_contribution, na.rm = TRUE)) %>%
    mutate(
      share = contrib / sum(contrib),
      context = "CH")
  bind_rows(to_all, to_ch)
}

pillar_area_lookup <- df %>%
  filter(framework_programme == "H2020") %>%
  count(Pillar = pillar, Area = thematic_priority_abbr) %>%
  filter(!is.na(Area)) %>%
  select(-n)

area_labels <- df %>%
  get_contrib_area() %>%
  filter(context == "CH") %>%
  left_join(pillar_area_lookup, by = "Area") %>%
  arrange(desc(contrib)) %>%
  slice_head(n = 20) %>%
  arrange(desc(Pillar), desc(contrib)) %>%
  select(Area)

contribs <- df %>%
  get_contrib_area() %>%
  select(-share) %>%
  mutate(
    contrib = round(1e-6 * contrib, 1),
    contrib = format(contrib, big.mark = "'", decimal.mark = ".", trim = TRUE))

pal <- c("#B91646", "#DFD8CA")

area_labels %>%
  left_join(get_contrib_area(df), by = "Area") %>%
  select(-contrib) %>%
  pivot_wider(names_from = context, values_from = share) %>%
  pivot_longer(
    cols = c("Total", "CH"), 
    names_to = "context", 
    values_to = "share") %>%
  left_join(contribs, by = c("Area", "context")) %>%
  mutate(
    context = str_replace(context, "CH", "Swiss Institutions"),
    context = str_replace(context, "Total", "All Institutions"),
    context = fct_rev(context),
    Area = str_remove(Area, "LEIT-"),
    Area = str_remove(Area, "INNOSUP"),
    Area = fct_inorder(Area),
    Area = fct_rev(Area)) %>%
  ggplot(aes(x = Area, y = share, fill = fct_rev(context))) +
  geom_bar(stat = "identity", width = .7, position = "dodge") +
  geom_text(
    aes(label = contrib),
    position = position_dodge(width = .9),
    hjust = -.25,
    size = 3.5,
    family = "Gill Sans"
  ) +
  coord_flip() +
  scale_fill_manual(values = rev(pal), guide = guide_legend(reverse = TRUE)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), expand = c(.1, 0)) +
  theme_minimal() +
  labs(
    title = "Funding by Programme Area Under H2020",
    subtitle = "(MioEUR)")+
  theme(
    text = element_text(size = 14, family = "Gill Sans", color = "gray10"),
    plot.title.position = "plot",
    legend.position = "top",
    legend.title = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_text(margin = margin(0, -25, 0, 0)))
```

### Figure 4 (Original)

<img src="https://www.sbfi.admin.ch/sbfi/en/home/research-and-innovation/international-cooperation-r-and-i/eu-framework-programmes-for-research/f-f-swiss-participation/_jcr_content/par/accordion_16708559/items/abbildung_4_beitr_ge/accordionpar/image_896504598/image.imagespooler.png/1629209007946/abb4_en.png" alt="SERI Figure 4. Funding under by programme area (CHF m), in relation to total funding" width="600"/>

Source: [Figure 4. Funding under by programme area (CHF m), in relation to total funding](https://www.sbfi.admin.ch/sbfi/en/home/research-and-innovation/international-cooperation-r-and-i/eu-framework-programmes-for-research/f-f-swiss-participation.html)


## Participation by Sector

### EU Contribution to Swiss Institutions in Research and HigherEd

```{r overview-in-research}
df %>%
  filter(country_code == "CH") %>%
  filter(org_type == "Research & Education") %>%
  get_participation()
```

### Typology of the Top-10 Participating Countries

```{r additional-context, fig.asp=1}
pal <- c(
  rep("#767676", ch_rank - 1),
  "#800000",
  rep("#767676", 10 - ch_rank))

## share to Pillar I
contribution_share_pillar <- df %>%
  filter(!is.na(pillar)) %>%
  semi_join(top_participants %>%
    select(country = Country) %>%
    head(10),
  by = "country"
  ) %>%
  group_by(
    Country = country,
    Pillar = pillar
  ) %>%
  summarise(`Contribution` = sum(eu_contribution)) %>%
  mutate(share_pillar = 100 * prop.table(Contribution)) %>%
  filter(Pillar == "I") %>%
  select(-Pillar, -Contribution)

## share to Resesearch and HigherEd
contribution_share_reshed <- df %>%
  filter(!is.na(pillar)) %>%
  semi_join(top_participants %>%
    select(country = Country) %>%
    head(10),
  by = "country"
  ) %>%
  group_by(
    Country = country,
    Type = org_type
  ) %>%
  summarise(`Contribution` = sum(eu_contribution)) %>%
  mutate(share_reshed = 100 * prop.table(Contribution)) %>%
  filter(Type == "Research & Education") %>%
  select(-Type, -Contribution)

top_participants %>%
  inner_join(contribution_share_reshed, by = c("Country")) %>%
  inner_join(contribution_share_pillar, by = c("Country")) %>%
  mutate(
    Country = fct_rev(fct_inorder(Country)),
    contrib_country = contrib_country / 1000000
  ) %>%
  ggplot(
    aes(
      x = share_pillar,
      y = share_reshed,
      label = Country,
      col = Country,
      size = contrib_country)) +
  geom_point(fill = NA) +
  geom_text_repel(
    size = 5,
    show.legend = FALSE,
    box.padding = .75,
    family = "Gill Sans") +
  guides(size = guide_legend(
    title = "Total EU Contribution (MioEUR) :",
    override.aes = list(fill = "gray45", col = "gray45", shape = 21))) +
  labs(
    x = glue("Contributions to Pillar I: ",
             "Excellent Science\n",
             "(% of the total contribution to the country)"),
    y = glue("Contribution to Institutions in Research and HigherEd\n",
             "(% of the total contribution to the country)")) +
  scale_color_manual(values = rev(pal), guide = "none") +
  theme_minimal() +
  theme(
    text = element_text(size = 16, family = "Gill Sans", color = "gray10"),
    plot.title.position = "plot",
    legend.position = "top")
```


### Relative Semester Distribution of EU Contribution

```{r, time-distribution, fig.asp=.6}
pal <- c("#800000", "#767676")

ch <- df %>%
  filter(!is.na(call_deadline_date)) %>%
  filter(!is.na(org_type)) %>%
  filter(country_code == "CH") %>%
  mutate(
    call_semester = floor_date(call_deadline_date, "6 months")) %>%
  group_by(Country = country, org_type, Semester = call_semester) %>%
  summarise(contrib = sum(eu_contribution, na.rm = TRUE)) %>%
  mutate(Share = 100 * prop.table(contrib)) %>%
  select(-contrib)

rest <- df %>%
  filter(!is.na(call_deadline_date)) %>%
  filter(!is.na(org_type)) %>%
  filter(!country_code %in% c("CH")) %>%
  mutate(
    call_semester = floor_date(call_deadline_date, "6 months")) %>%
  group_by(org_type, Semester = call_semester) %>%
  summarise(contrib = sum(eu_contribution)) %>%
  mutate(Share = 100 * prop.table(contrib)) %>%
  select(-contrib) %>%
  mutate(Country = "Remaining Countries") %>%
  relocate(Country)

ch %>%
  bind_rows(rest) %>%
  mutate(Country = factor(
    Country,
    levels = c("Switzerland", "Remaining")
  )) %>%
  ggplot(aes(x = Semester, y = Share, fill = Country)) +
  geom_bar(
    stat = "identity",
    width = 120,
    position = position_dodge(width = 128)) +
  labs(
    title = "Time Distribution of EU Contribution by Institution Sector",
    subtitle = glue(
      "<span style='color:{pal[1]};'>Switzerland</span> ",
      "vs <span style='color:{pal[2]};'>Remaining Countries</span>"
    ),
    x = "Call Semester",
    y = "%"
  ) +
  scale_fill_manual(values = pal, guide = "none") +
  facet_wrap(org_type ~ ., ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(
    plot.title.position = "plot",
    plot.subtitle = element_markdown(),
    text = element_text(
      size = 14,
      family = "Gill Sans",
      color = "gray10"
    )
  )
```

---

## Annex

### Top Recipients

#### Top 30 in Research and Higher Education

```{r}
df %>%
  filter(country_code == "CH") %>%
  filter(org_type == "Research & Education") %>%
  group_by(Institution = legal_name) %>%
  summarise(
    `Nb Projects` = n_distinct(project_id),
    `EU Contribution` = sum(eu_contribution, na.rm = TRUE)
  ) %>%
  arrange(desc(`EU Contribution`)) %>%
  mutate(Institution = str_to_title(Institution)) %>%
  slice_head(n = 30) %>%
  reactable(
    columns = list(
      `Nb Projects` = colDef(sortable = TRUE),
      `EU Contribution` = colDef(
        sortable = TRUE,
        format = colFormat(
          currency = "EUR",
          separators = TRUE,
          digits = 2
        )
      )
    ),
    highlight = TRUE,
    searchable = TRUE,
    sortable = FALSE,
    showSortable = TRUE,
    compact = TRUE,
    minRows = 10
  )
```

#### Top 30 not in Research and Higher Education

```{r}
df %>%
  filter(country_code == "CH") %>%
  # remove international organisations
  filter(org_type == "Others (Excl. R&Ed)") %>%
  group_by(Institution = legal_name) %>%
  summarise(
    `Nb Projects` = n_distinct(project_id),
    `EU Contribution` = sum(eu_contribution, na.rm = TRUE)
  ) %>%
  arrange(desc(`EU Contribution`)) %>%
  mutate(Institution = str_to_title(Institution)) %>%
  slice_head(n = 30) %>%
  reactable(
    columns = list(
      `Nb Projects` = colDef(sortable = TRUE),
      `EU Contribution` = colDef(
        sortable = TRUE,
        format = colFormat(
          currency = "EUR",
          separators = TRUE,
          digits = 2
        )
      )
    ),
    highlight = TRUE,
    searchable = TRUE,
    sortable = FALSE,
    showSortable = TRUE,
    compact = TRUE,
    minRows = 10
  )
```


```{r visual-overview, eval=FALSE}
# Differentiate pillars 1 to III
stats_depth_1 <- df %>%
  filter(!is.na(pillar)) %>%
  filter(country_code == "CH") %>%
  filter(!pillar %in% chr(roman(1:3))) %>%
  mutate(pillar = fct_rev(pillar)) %>%
  group_by(pillar, pillar_descr) %>%
  summarise(total_eu_contrib = sum(eu_contribution)) %>%
  ungroup() %>%
  # for printing
  mutate(
    pillar_descr = ifelse(
      chr(pillar) %in% chr(roman(4:5)),
      str_c(pillar, ". ", pillar_descr),
      chr(pillar)
    ),
    thematic_priority_abbr = NA
  ) %>%
  # for binding
  select(
    pillar = pillar_descr,
    thematic_priority = thematic_priority_abbr,
    contrib = total_eu_contrib
  )

stats_depth_2 <- df %>%
  filter(!is.na(pillar)) %>%
  filter(country_code == "CH") %>%
  filter(pillar %in% chr(roman(1:3))) %>%
  mutate(pillar = fct_rev(pillar)) %>%
  group_by(pillar, pillar_descr, thematic_priority_abbr) %>%
  summarise(total_eu_contrib = sum(eu_contribution)) %>%
  ungroup() %>%
  # for printing
  unite("pillar", starts_with("pillar"), sep = ". ") %>%
  # for binding
  select(pillar,
    thematic_priority = thematic_priority_abbr,
    contrib = total_eu_contrib
  )

stats_depth_2 %>%
  bind_rows(stats_depth_1) %>%
  mutate(
    thematic_priority = replace_na(thematic_priority, ""),
    contrib_txt = contrib / 1000000,
    contrib_txt = format(
      contrib_txt,
      big.mark = ",",
      decimal.mark = ".",
      trim = TRUE,
      digits = 2
    ),
    contrib_txt = str_c("M€", contrib_txt)
  ) %>%
  write_csv(here("data", "h2020-treemap.csv"))
# plot with https://rawgraphs.io
```

```{r ch-type-role, fig.asp=.7, eval = FALSE}
pal <- c("#B91646", "#DFD8CA")
df %>%
  filter(!is.na(pillar)) %>%
  filter(country_code == "CH") %>%
  mutate(
    pillar = fct_other(pillar, keep = chr(roman(1:5))),
    pillar = fct_relevel(pillar, "Other", after = 0)
  ) %>%
  group_by(pillar, org_type, role = partner_role) %>%
  summarise(contrib = sum(eu_contribution, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(org_type = fct_rev(org_type)) %>%
  arrange(pillar, org_type) %>%
  ggplot(aes(x = pillar, y = contrib, fill = org_type)) +
  geom_col(position = position_stack(reverse = FALSE)) +
  scale_fill_manual(values = rev(pal), guide = guide_legend(reverse = TRUE)) +
  scale_y_continuous(labels = unit_format(
    suffix = "",
    scale = 1e-6,
    big.mark = ","
  )) +
  coord_flip() +
  facet_wrap(role ~ ., ncol = 1) +
  theme_minimal() +
  labs(
    title = "Distribution of EU Contribution to Swiss Institutions",
    subtitle = "By Pillar, Institution Sector, and Instituion Role",
    x = "Pillar",
    y = "EU Contribution to Swiss Institutions (MioEUR)"
  ) +
  theme(
    text = element_text(size = 15, family = "Gill Sans", color = "gray10"),
    plot.title.position = "plot",
    legend.position = "top",
    legend.title = element_blank(),
    axis.title.y = element_text(
      angle = 0, 
      vjust = 1.05, 
      margin = margin(r = -30))
  )

```
