---
title: "Swiss Participation in the EU R&I Programmes"
output:
  html_document:
    toc: false
---

```{r, include=FALSE, results='asis'}
library(knitr)
library(metathis)

opts_chunk$set(echo = FALSE)
meta() %>%
  meta_description(
    "Swiss Participation in FP7 and Horizon 2020") %>%
  meta_name("github-repo" = "zambujo/swissparticipation") %>%
  meta_viewport() %>%
  meta_social(
    title = "Switzerland in FP7&H2020",
    url = "https://zambujo.github.io/swissparticipation",
    image = "https://zambujo.github.io/assets/images/treemap.jpg",
    image_alt = "CH H2020 Participation",
    og_type = "website",
    og_author = "João Martins",
    twitter_card_type = "summary",
    twitter_creator = "@whereisjoao"
  )
```

<!-- badges: start -->
[![](https://img.shields.io/github/watchers/zambujo/swissparticipation?style=social)](https://github.com/zambujo/swissparticipation)
[![](https://img.shields.io/github/issues/zambujo/swissparticipation)](https://github.com/zambujo/swissparticipation/issues)
<!-- badges: end -->

<br />

```{r motivation, child=here::here("Rmd", "motivation_en.rmd")}
```

<br />

## Background

```{r background, child=here::here("Rmd", "background_en.rmd")}
```

## Swiss Participation Over Time {.tabset .tabset-fade .tabset-pills}

```{r swiss-participation, child=here::here("Rmd", "swiss-participation_en.rmd")}
```

### Contributions

```{r semester-contributions}
include_graphics("https://zambujo.github.io/swissparticipation/seri-figures_files/figure-html/swiss-countributions-1.svg")
```

### Participation

```{r}
include_graphics("https://zambujo.github.io/swissparticipation/seri-figures_files/figure-html/swiss-participations-1.svg")
```

### Coordinations

```{r}
include_graphics("https://zambujo.github.io/swissparticipation/seri-figures_files/figure-html/coordinations-overviews-1.svg")
```

## Top Participating Countries {.tabset .tabset-fade .tabset-pills}

```{r top-participants, child=here::here("Rmd", "top-participants_en.rmd")}
```

### FP7 (2007-2013)

```{r}
include_graphics("https://zambujo.github.io/swissparticipation/seri-figures_files/figure-html/visual-countries-fp7-1.svg")
```

### H2020 (2014-2020)

```{r}
include_graphics("https://zambujo.github.io/swissparticipation/seri-figures_files/figure-html/visual-countries-1.svg")
```

##  H2020 structure {.tabset .tabset-fade .tabset-pills}

```{r h2020-structure, child=here::here("Rmd", "h2020-structure_en.rmd")}
```

### All countries

```{r}
include_graphics("https://raw.githubusercontent.com/zambujo/swissparticipation/main/data/treemap-h2020-all.svg")
```

### Switzerland

```{r}
include_graphics("https://raw.githubusercontent.com/zambujo/swissparticipation/main/data/treemap-h2020-ch.svg")
```

## Participation profile {.tabset .tabset-fade .tabset-pills}

```{r participation-profile, child=here::here("Rmd", "participation-profile_en.rmd")}
```

### Excellent Science/Research and HigherEd

```{r}
include_graphics("https://zambujo.github.io/swissparticipation/swiss-participation_files/figure-html/country-profile-1-1.svg")
```

### Industrial Leadership/Private Sector

```{r}
include_graphics("https://zambujo.github.io/swissparticipation/swiss-participation_files/figure-html/country-profile-2-1.svg")
```

## Outlook

```{r outlook, child=here::here("Rmd", "outlook_en.rmd")}
```

```{r}
include_graphics("https://raw.githubusercontent.com/zambujo/swissparticipation/main/data/treemap-he.svg")
```

---

## Varia

```{r see-also, child=here::here("Rmd", "varia_en.rmd")}
```
